The Diminishing Returns of Music in the Digital Era

### 2024-04-17 paper/paper.qmd
results section done

library(tidyverse)
library(knitr)
library(readr)
library(ggplot2)
library(GGally)
library(palmerpenguins)
```

@@

## Source {#sec-source}
The main data source is a dataset from the Recording Industry Association of America (RIAA) (@RIAA), tracking recorded music revenues by format in the United States from 1973 to 2019. This dataset was generated by Matt Bass and has been adjusted for inflation, offering a standardized monetary comparison across five decades of music industry economics. Since the RIAA have yet to properly respond to requests for the use of their dataset, we made use of a Kaggle dataset from Kaggle user "Larxel" that cites the RIAA dataset.

@@

Each category represents a particular mode of music consumption, and the revenue figures associated with these formats are used to track shifts in consumer preferences and technological advancements. All physical formats and digital formats were cleaned and grouped together.

@@

The RIAA dataset published by @RIAA was cleaned to account for the irrelevence of the "number_of_records" column. Using the dataset, we were able to replicate all of the means and standard deviations of the music formats, as shown in @tbl-summary1. The summary statistics illustrate how the high standard deviations across the board reflect a wide variation in yearly revenue, which could be attributed to a certain format's lifecycle or other market changes over time. For example, 8-Track tapes, despite their irrelevance, have an intriguingly high mean revenue.

@@

@fig-revenuechange illustrates the year-over-year percentage change in revenue for various music formats, demonstrating the music industry's dynamic nature over time. For example, the sharp volatility observed in the cassette format's line is indicative of its rise and fall in popularity, reflective of technological obsolescence and subsequent market shifts. CDs exhibit less volatility but show a general decline in recent years, which correlates with the proliferation of digital music. The download single format and LP/EP show resurgence possibly due to niche market appeal and the revival of vinyl records among collectors and other enthusiasts. These revivals can be traced to social media trends as well, further propagating the internet's immense influence on the music industry. The paid subscription format presents a growing trend, emphasizing the industry's shift towards streaming services as the primary source of music consumption. This graph not only highlights broader trends in the industry, but it also lines up the death of physical media with the rise of the internet, which have reshaped the way music is distributed and monetized. The main concern of the contemporary music industry is also present in this graph as the most consumed format of music (paid subscriptions) are declining in revenue year by year.

@@

ggplot(top_formats_data, aes(x = year, y = yoy_change, color = format)) +
  geom_line() +
  scale_y_continuous(limits = c(-100, 100),
                     labels = scales::percent) +
  labs(title = "Year-Over-Year Revenue Change for Top Music Formats",
       x = "Year",
       y = "Percentage Change",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@@

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

@@

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

@@

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

@@

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

>>>>

library(tidyverse)
library(knitr)
library(readr)
library(ggplot2)
library(brms)
library(palmerpenguins)
```

@@

## Source {#sec-source}
The main data source is a dataset from the Recording Industry Association of America (RIAA) (@RIAA), tracking recorded music revenues by format in the United States from 1973 to 2019. This dataset was generated by Matt Bass and has been adjusted for inflation, offering a standardized monetary comparison across five decades of music industry economics. Since the RIAA have yet to properly respond to requests for the use of their dataset, we made use of a Kaggle dataset from Kaggle user "Larxel" that cites the RIAA dataset. There could have been other datasets from Kaggle used but they weren't because they didn't cite the RIAA dataset.

@@

Each category represents a particular mode of music consumption, and the revenue figures associated with these formats are used to track shifts in consumer preferences and technological advancements.

@@

The RIAA dataset published by @RIAA was cleaned to account for the irrelevence of the "number_of_records" and "metric" columns. Using the dataset, we were able to replicate all of the means and standard deviations of the music formats, as shown in @tbl-summary1. The summary statistics illustrate how the high standard deviations across the board reflect a wide variation in yearly revenue, which could be attributed to a certain format's lifecycle or other market changes over time. For example, 8-Track tapes, despite their irrelevance, have an intriguingly high mean revenue.

@@

@fig-revenuedistribution presents a vivid picture of how music format revenues have shifted over approximately four decades and the rise and fall of various music formats, echoing the technological and cultural shifts over time. Notably, the graph illustrates the dominance of CDs in the 90s through early 2000s, highlighting the era's preference for physical media. The subsequent decline of CDs and rise of digital formats reflect the impact of the internet and mobile technology on music consumption. The recent surge in streaming services, visible in the graph, represents the latest phase in music distribution, where conveniency takes precedence over physical ownership. This shift not only marks a change in consumer habits but also has broad implications for revenue models in the music industry, affecting everything from artist compensation to the survival of traditional record stores. The data in this graph tells the story of an industry that has had to continuously adapt to the rapid evolution of technology and consumer preferences.

```{r}
#| echo: false
#| eval: true
#| label: fig-revenuedistribution
#| fig-cap: Revenue Distribution by Year and Format
#| warning: false
#| results: asis

@@

yearly_format_revenue <- music_data %>%
  group_by(year, format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  ungroup()

ggplot(yearly_format_revenue, aes(x = year, y = total_revenue, fill = format)) +
  geom_bar(stat = "identity", position = "stack") +
  theme_minimal() +
  labs(x = "Year", y = "Total Revenue ($)", fill = "Format") +
  theme(legend.position = "bottom")
```

@@

## Revenue Percent Change in the Digital Age

ggplot(top_formats_data, aes(x = year, y = yoy_change, color = format)) +
  geom_line() +
  scale_y_continuous(limits = c(-100, 100),
                     labels = scales::percent) +
  labs(
       x = "Year",
       y = "Percentage Change",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

@fig-revenuechange illustrates the year-over-year percentage change in revenue for various music formats, demonstrating the music industry's dynamic nature over time. For example, the gradual rise and later steep decline observed in the cassette format's line is indicative of its rise and fall in popularity, reflective of technological obsolescence and subsequent market shifts. CDs exhibit a general decline in recent years, which correlates with the proliferation of digital music. The download single format and LP/EP show resurgence possibly due to niche market appeal and the revival of vinyl records among collectors and other enthusiasts. These revivals can be traced to social media trends as well, further propagating the internet's immense influence on the music industry. The paid subscription format presents a growing trend, emphasizing the industry's shift towards streaming services as the primary source of music consumption. This graph not only highlights broader trends in the industry, but it also lines up the death of physical media with the rise of the internet, which have reshaped the way music is distributed and monetized.

## Volatile Revenue Trends

@fig-revenuetrends illustrates the revenue trends over time for various music formats, demonstrating the music industry's sharp volatility regarding revenue over time. The graph generally indicates that the more volatile the revenue for a given music format, the more revenue it will generate. However, the more consistent the revenue for a given music format, the less revenue it will generate. This also demonstrates the revenue trend between physical and digital sales. Physical sales generate more revenue but are less consistent whereas digital sales generate less revenue but are more consistent. This could be due to the convenience that the internet and the ever-increasingly connected world provide in regards to music consumption. Additionally, the graph shows a general decline in recent years for physical music formats such as CDs, which further correlates with the proliferation of digital music and the internet. This is the best illustration of the main concern of the contemporary music industry where the most prominent format of music (paid subscriptions) is failing to generate as much revenue as it seemingly should.

```{r}
#| echo: false
#| eval: true
#| label: fig-revenuetrends
#| fig-cap: Revenue Trends per Music Format
#| warning: false
#| results: asis

top_formats <- music_data %>%
  group_by(format) %>%
  summarise(total_revenue = sum(value_actual, na.rm = TRUE)) %>%
  top_n(7, total_revenue) %>%
  pull(format)

music_data_top <- music_data %>%
  filter(format %in% top_formats)

ggplot(music_data_top, aes(x = year, y = value_actual, color = format, group = format)) +
  geom_line() +
  scale_y_continuous(labels = scales::comma) + # Use scales package to format y axis
  labs(
       x = "Year",
       y = "Total Revenue",
       color = "Format") +
  theme_minimal() +
  theme(legend.position = "bottom")
```

## Model Results

@@

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false
#| results: asis

@@

bayesian_model <- brm(value_actual ~ format, data = music_data, 
                      family = gaussian(), 
                      prior = c(set_prior("normal(0, 10)", class = "b"),
                                set_prior("cauchy(0, 2)", class = "sigma")),
                      chains = 4, iter = 2000, warmup = 500, 
                      control = list(adapt_delta = 0.95),
                      refresh = 0)
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| message: false
#| warning: false
#| results: asis

modelsummary::modelsummary(
  list("Bayesian Linear Regression" = bayesian_model),
  output = "markdown",
  fmt = 2
)
```

@tbl-modelresults present the results of the Bayesian linear regression analysis, examining the relationship between music formats and revenue generation. The model's coefficients (b) for various formats—such as cassette, CD, and different digital formats—are listed alongside their numerical estimates, which represent the expected change in revenue generation associated with each format, holding all else equal.

The coefficient for the intercept indicates the baseline revenue when all other predictors are at zero. The significant figure of 761.20 is the baseline revenue for the reference category of music formats. Negative coefficients, like those for paid subscriptions and downloads, suggest that these formats are associated with a decrease in revenue relative to the reference category. In contrast, positive values, like the one for vinyls, imply a potential increase in revenue.

Overall, this table provides a comprehensive overview of how different music formats are estimated to impact revenue generation within the scope of the data analyzed, offering valuable insights for stakeholders in the music industry regarding which formats may be more financially viable or are declining in terms of revenue contribution.

@@

# ```{r}
# #| eval: true
# #| echo: false
# #| message: false
# #| warning: false
# #| label: fig-ppcheckandposteriorvsprior
# #| layout-ncol: 2
# #| fig-cap: "Examining how the model fits, and is affected by, the data"
# #| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]
# 
# pp_check(first_model) +
#   theme_classic() +
#   theme(legend.position = "bottom")
# 
# posterior_vs_prior(first_model) +
#   theme_minimal() +
#   scale_color_brewer(palette = "Set1") +
#   theme(legend.position = "bottom") +
#   coord_flip()
# ```

@@

# ```{r}
# #| echo: false
# #| eval: true
# #| message: false
# #| warning: false
# #| label: fig-stanareyouokay
# #| fig-cap: "Checking the convergence of the MCMC algorithm"
# #| fig-subcap: ["Trace plot", "Rhat plot"]
# #| layout-ncol: 2
# 
# plot(first_model, "trace")
# 
# plot(first_model, "rhat")
# ```
